---
title: "Determination of oxygen mass transfer coefficient IKLa) in batch reactor"
author: "Rafael Ventura"
date: "23 de junio de 2017"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

## Objectives

## Background
Assignment is based on experiment for measuring D.O.
````{r dataset, echo=FALSE}
library(ggplot2)
library(readxl)
dataset <- read_excel("D:/Dropbox/R/projects/water-sanitation-modelling/Data.xls", 
     sheet = "Data_all_and_calcul (2)", skip = 5)
ggplot(dataset, aes(time_hhmmss,doc_mgl,fill=phase))+geom_line(aes(colour=phase, group=phase))
```

## Procedure

Dissolved Oxygen concentration at equilibrium will be calculated using the Henry's Law. For that purpose:
first, determine the barometrice pressure at 2600 [m], given P0, T, R, g, M using the barometric formula:

(@) $P=P0*e^\frac{-Mgz}{RT}$
where:

$M=1,785*10^20$ [mol]; $P0=101325$ [Pa]; $T=298,15$ [K]; $R=3,314$ [J*K/mol]; $g=9,80$ [m/s^2]

$P = 101325*exp(-\frac{28,96*10^-3*9,8*2600}{8,314*298,15}) = 75214$ [Pa]

Second, determine dissolve oxygen concentration:

$SO_eq = \frac {P}{K_{H}}$

where $P= 75214$ [Pa]

$K_{H} = 78,020*10^3$ [Pa*L/mol] ^[Sanders, R (2015), "Compilation of Henry's Law constants for water as a solvent". Atmos. Chem. Phys 15: 4399-4981]

Partial pressure for oxygen is $P_{O2}=0,21*P = 15795$ [Pa]

$SO_eq = \frac {15795*32*1000}{78,020*10^3} = 6.4$ [mg/L]

As substrate is depleted, Biomass undergoes endogenous respiration process in phase 1. As stated in problem definition, OUR is constant and it is calculated by mean of the following expression:

$OUR = -\frac{DO_{end-of-phase1}-DO_{t0}}{t_{end-of-phase1}-t_{0}}$ [mg/L-h]

```{r OUR1, echo=FALSE}
library(readxl)
library(lattice)
dataset <- read_excel("D:/Dropbox/R/projects/water-sanitation-modelling/Data.xls", 
     sheet = "Data_all_and_calcul (2)", skip = 5)
p1_dataset <- subset(dataset,dataset$phase==1)
OURdC <- (p1_dataset[nrow(p1_dataset),"doc_mgl"]-p1_dataset[1,"doc_mgl"])
OURdt <- (as.numeric(p1_dataset[nrow(p1_dataset),"time_hhmmss"]-p1_dataset[1,"time_hhmmss"])/60)
print(paste("OUR_OHO = ", as.character(round(-OURdC/OURdt,2))," [mg/L-h]", ""))

```
Alternatively, $OUR$ can be obtained from linear regression of dissolved oxygen vs time
```{r OUR2, echo=FALSE}
library(readxl)
library(lattice)
dataset <- read_excel("D:/Dropbox/R/projects/water-sanitation-modelling/Data.xls", 
     sheet = "Data_all_and_calcul (2)", skip = 5)
p1_dataset <- subset(dataset,dataset$phase==1)
p1_dataset$time_hh <- as.numeric(p1_dataset$time_hhmmss- as.POSIXct("1899-12-31 14:51:40"))/60
OUR_lm <- lm(doc_mgl ~ time_hh, data = p1_dataset)
print(OUR_lm)
xyplot(doc_mgl ~ time_hh, data = p1_dataset, panel = function(x, y, ...) {
                 panel.xyplot(x, y, ...)
                 panel.lmline(x, y, ...)
               })
print(paste("OUR_OHO = ", as.character(OUR_lm$coefficients[2])," [mg/L-h]", ""))
```
$K_{La}$ can obtained directly without need to use AQUASIM, by plotting $ln(\frac{SO_{eq}-SO}{SO_{eq}-SO_{0}})$ vs time plus linear regression to obtain the linear slope coefficient as $K_{La}$ with the assumption reactor is modelled as CSTR. 
```{r Kal, echo=FALSE }
library(lattice)
library(readxl)
dataset <- read_excel("D:/Dropbox/R/projects/water-sanitation-modelling/Data.xls", 
     sheet = "Data_all_and_calcul (2)", skip = 5)
p2_dataset <- subset(dataset,dataset$phase==2)
p2_dataset$time_hh <- as.numeric(p2_dataset$time_hhmmss- as.POSIXct("1899-12-31 14:51:40"))/60-0.3877778
# print(p2_dataset$time_hh[nrow(p2_dataset)])
p2_dataset$LN <- log((6.4-p2_dataset$doc_mgl)/(6.4-p2_dataset$doc_mgl[1]))
p2_dataset_lm <- subset(p2_dataset, p2_dataset$time_hh < 0.15)
Kla_lm <- lm(LN ~ time_hh, data = p2_dataset_lm)
print(Kla_lm)
xyplot(LN ~ time_hh, data = p2_dataset_lm, panel = function(x, y, ...) {
                 panel.xyplot(x, y, ...)
                 panel.lmline(x, y, ...)
               })
print(paste("KLa = ", as.character(round(Kla_lm$coefficients[2],3))," [1/h]", ""))


```
**Simulation with AQUASIM**

State variable is dissolved oxygen concentration (SO) in [mg/L], read from a .CSV file.

Process variable is time in [h].

Process is defined as reaction in CSTR

Reaction rate is defined as $K_{La}*(SO_{sat} - SO)$, where KLa and SOsat are set to be -6.479 [1/h] and 6.4 [mg/L], respectively as per estimated values calculated above. Reactor volumen is set to be 1,64 [L].

Neither diffusive nor adventive link are considered.

Time interval simulation is set to be 0.10 [h] from 0 to 0.3902778.

Simulation is run obtaining the following output shown in figure 2 

## Discussion

## Conclussions

